---
title: "Calculate Stabilization Rates for Pitcher Stats"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
  word_document: default
---

We will use the database created in the Read Pitch-By-Pitch Data.Rmd file to calculate stabilization rates for various pitcher stats.

An explanation of stabilization rate is found in this article, which calculates the stabilization rate of many baseball stats:
https://blogs.fangraphs.com/a-long-needed-update-on-reliability/

We will calculate stabilization using the same methodology, reproducing the study in the article over a larger sample, from 2008 to 2019.

We will focus on the following stats: 

  strikeout percentage - the percent of a pitcher's plate appearances that end in a strikeout
  
  swinging strike percentage - the percent of a pitcher's pitches that end with a swing and miss
  
  called strike percentage - the percent of a pitcher's pitches that end with a called strike
  
  foul percentage - the percent of a pitcher's pitches that end with a foul
  
  ball percentage - the percent of a pitcher's pitches that end with a ball
  
  hit by pitch percentage - the percent of a pitcher's pitches that end with a hit by pitch
  
  in play percentage - the percent of a pitcher's pitches that are put in play by the hitter
  
  pitches per plate appearance - the average number of pitches a pitcher throws in each plate appearance

First, load the packages we will need:
```{r}

library(dplyr)
library(RSQLite)
library(data.table)
library(ltm)
library(ggplot2)
library(ggthemes)
library(plotly)

```

Next, get the data we need from our database
```{r}

#load in the columns we will use to calculate our stabilization rates

#the 'numPitch' column is added to count up the number of pitches each pitcher throws in a given season
#this will make taking samples of the data much easier
#we will filter out intentional balls (which are decided by the manager and not indicative of a pitcher's skill) and bunts (which have much more to do with the hitter's skill than the pitcher's, and might skew the data)
pitches <- as.data.frame(dbGetQuery(mySQLiteDB, "SELECT game_year, game_month, game_day,
		player_name, pitcher, events, description, pitch_type, release_speed, plate_x, plate_z,
		p_throws, type, balls, strikes, pitch_number, at_bat_number,
		
		RANK() OVER(
				PARTITION BY game_year, pitcher
				ORDER BY game_month ASC, game_day ASC, at_bat_number ASC, pitch_number ASC) numPitch

FROM PitchByPitch
WHERE description NOT IN ('intent_ball', 'missed_bunt', 'foul_bunt', 'bunt_foul_tip')"))

#disconnect from SQLite
dbDisconnect(mySQLiteDB)

```

We will clean up our new 'pitches' dataframe by getting rid of plate appearances that do not reflect a pitcher's skill, such as catcher's interferences and sacrifice bunts
```{r}

#change NA values in events to 'None'
pitches$events[is.na(pitches$events)] <- 'None'

#filter out pitches we don't want
pitches <- pitches %>% filter(events != 'sac_bunt' & events != 'sac_bunt_double_play' & events != 'catcher_interf')

```

We will also add a column to 'pitches' that calculates the total number of pitches a pitcher threw in a given season:
```{r}

numPit <- pitches %>% group_by(pitcher, game_year) %>% 
  summarise(totPit = n())
pitches <- pitches %>% left_join(numPit)

```

Next, create a separate dataframe that only contains pitches that ended a plate appearance. There will be exactly one entry in this dataframe for each plate appearance in the dataset.
```{r}

#Filter the 'pitches' dataframe to only contain pitches that ended a plate appearance
PAs <- pitches %>% filter(events %in% c('field_out', 'single', 'strikeout', 'force_out', 'walk', 'double', 'field_error', 'sac_fly', 'home_run', 'double_play', 'hit_by_pitch', 'grounded_into_double_play', 'triple', 'fielders_choice', 'strikeout_double_play', 'fielders_choice_out', 'sac_fly_double_play', 'triple_play'))

#like we created the 'totPit' column to calculate the total number of pitches a pitcher had in a given season, we will do the same with plate appearances
numPAs <- PAs %>% group_by(pitcher, game_year) %>% 
  summarise(totPAs = n())
PAs <- PAs %>% left_join(numPAs)

#add a unique index for each row in PAs (this will make taking samples easier)
PAs <- PAs %>% mutate(PA_index = row_number())

```

Next, we will create functions to perform each of the operations we need to perform in the analysis.

```{r}

#the alphaCalc function will calculate the cronbach's alpha value (a way to measure the reliability of a statistic in a sample of values, what we want to measure for our pitcher statistics)
#an explanation for using cronbach's alpha can be found here:
#https://blogs.fangraphs.com/a-new-way-to-look-at-sample-size-math-supplement/

#the function takes in a sample size, a metric function (which will be defined below), and a denominator (either Pitches or PAs, for plate appearances)
alphaCalc <- function(ss, metricFUN, denom){
  
  #make sure we get the same sample each time we run the code
  set.seed(1234)
  
  #extract the statistic's name from the 'metricFUN' argument to the function
  metric <- gsub("Calc", "", metricFUN)
  
  #we will perform different tasks depending on the denominator
  
  #for denom = PAs:
  if(denom == 'PAs'){
    
      #get a sample of plate appearances (based on the 'ss' argument to the function) from each pitcher season
      selectSample <- PAs %>% group_by(pitcher, game_year) %>%
        filter(totPAs >= ss) %>%
        slice_sample(n = ss, replace = FALSE)
      
      #make a column to count up each PA for each pitcher season in the sample (there should be 'ss' number of PA for each pitcher season)
      selectSample <- selectSample %>% group_by(pitcher, game_year) %>% 
        mutate(pitcherPA = row_number())
      
      #calculate the stat we want (the stat saved in 'metric') using the 'metricFUN' argument passed into the function
      selectSample <- get(metricFUN)(selectSample) %>% dplyr::select(pitcher, game_year, pitcherPA, metric)
    
    #get the df in a form that we can measure the alpha
    selectSampleDT <- dcast(setDT(selectSample), pitcher + game_year ~ pitcherPA, value.var = metric)
    
    
    #for denom = Pitches
  } else if (denom == 'Pitches') {
    
    #get a sample (based on the 'ss' argument in the function) from each pitcher season
    selectSample <- pitches %>% group_by(pitcher, game_year) %>%
      filter(totPit >= ss) %>%
      slice_sample(n = ss, replace = FALSE)
    
    #make a column to count up each pitch for each pitcher season in the sample (there will be 'ss' number of pitches for each pitcher season)
    selectSample <- selectSample %>% group_by(pitcher, game_year) %>% 
      mutate(pitcherPit = row_number())
    
    #calculate the stat (saved in 'metric') we want using the 'metricFUN' argument passed into the function
    selectSample <- get(metricFUN)(selectSample) %>%
      dplyr::select(pitcher, game_year, pitcherPit, metric)
    
    #get the df in a form that we can measure the alpha
    selectSampleDT <- dcast(setDT(selectSample), pitcher + game_year ~
                              pitcherPit, value.var = metric)
    
    #if 'denom' doesn't equal PAs or Pitches, return an error message
  } else {
    
    return('Error: Invalid denominator')
    
  }
  
  #we will select only the columns of 'selectSampleDT' (our sample of plate appearances or pitches) we will need to calculate the alpha value
  data <- selectSampleDT[,3:(ss + 2)]
  
  #save the cronbach's alpha value in 'alpha_val'
  alpha_val <- cronbach.alpha(data)$alpha
  
  #return this variable
  return(alpha_val)
  
}

#the following functions calculate the different statistics we will be studying

#the function names can be passed into the 'metricFUN' argument in 'alphaCalc'
#for example, to calculate the alpha value of strikeout percentage for a sample size of 50, we can call:
#alphaCalc(50, 'K_pctCalc', 'PAs')

#This funciton calculates the strikeout percentage for a sample of plate appearances
K_pctCalc <- function(selectSample) {
  
  selectSample <- selectSample %>% mutate(K_pct = ifelse(events == 'strikeout', 1, 0))
  return(selectSample)
  
}

#This funciton calculates the average pitches per plate appearance for a sample of plate appearances
pitPerPACalc <- function(selectSample) {
  
  selectSample <- selectSample %>% rename(pitPerPA = pitch_number)
  return(selectSample)
  
}

#This funciton calculates the swinging strike percentage for a sample of pitches
swStr_pctCalc <- function(selectSample) {
  
  selectSample <- selectSample %>% 
    mutate(swStr_pct = ifelse(description %in% c('swinging_strike', 'swinging_strike_blocked',
                                                 'foul_tip', 'swinging_pitchout'), 1, 0))
  return(selectSample)
  
}

#This funciton calculates the called strike percentage for a sample of pitches
cStr_pctCalc <- function(selectSample) {
  
  selectSample <- selectSample %>% 
    mutate(cStr_pct = ifelse(description == 'called_strike', 1, 0))
  return(selectSample)
  
}

#This funciton calculates the foul percentage for a sample of pitches
foul_pctCalc <- function(selectSample) {
  
  selectSample <- selectSample %>% 
    mutate(foul_pct = ifelse(description %in% c('foul', 'foul_pitchout'), 1, 0))
  return(selectSample)
  
}

#This funciton calculates the ball percentage for a sample of pitches
ball_pctCalc <- function(selectSample) {
  
  selectSample <- selectSample %>% 
    mutate(ball_pct = ifelse(type == 'B' & description != 'hit_by_pitch', 1, 0))
  return(selectSample)
  
}

#This funciton calculates the in play percentage for a sample of pitches
inPlay_pctCalc <- function(selectSample) {
  
  selectSample <- selectSample %>% 
    mutate(inPlay_pct = ifelse(description == 'hit_into_play', 1, 0))
  return(selectSample)
  
}

#This funciton calculates the hit by pitch percentage for a sample of pitches
hbp_pctCalc <- function(selectSample) {
  
  selectSample <- selectSample %>% 
    mutate(hbp_pct = ifelse(description == 'hit_by_pitch', 1, 0))
  return(selectSample)
  
}

```

We will plot the alpha values of the various stats at different sample sizes. To do this, we need to calcualte the alpha values at these sample sizes. We can save them each in a list.

```{r}

#first, we will use a denominator of plate appearances and calculate strikeout percentage and pitches per plate appearance

#we will use sample sizes ranging from 5 plate appearances to 750 plate appearances
PAsampleSize <- seq(5, 750, 5)

#strikeout percentage
alphasK_pct <- lapply(PAsampleSize, alphaCalc, metricFUN = 'K_pctCalc', denom = 'PAs')

#pitches per plate appearance
alphaspitPerPA <- lapply(PAsampleSize, alphaCalc, metricFUN = 'pitPerPACalc', denom = 'PAs')

#make a dataframe with these values (this will allow us to plot the data using ggplot)
#the first column of the dataframe contains the sample size, the other two contain the alpha values for strikeout percentage and pitches per plate appearance
PA_alphas <- do.call(rbind, Map(data.frame, sampleSize = PAsampleSize, K_pctAlpha = alphasK_pct, pitPerPAAlpha = alphaspitPerPA))

#next, we will use a denominator of pitches and calculate the rest of our statistics

#we will use sample sizes ranging from 10 pitches to 2000 pitches
PitsampleSize <- seq(10, 2000, 10)

#swinging strike percentage
alphasswStr_pct <- lapply(PitsampleSize, alphaCalc, metricFUN = 'swStr_pctCalc', denom = 'Pitches')

#called strike percentage
alphascStr_pct <- lapply(PitsampleSize, alphaCalc, metricFUN = 'cStr_pctCalc', denom = 'Pitches')

#foul percentage
alphasfoul_pct <- lapply(PitsampleSize, alphaCalc, metricFUN = 'foul_pctCalc', denom = 'Pitches')

#ball percentage
alphasball_pct <- lapply(PitsampleSize, alphaCalc, metricFUN = 'ball_pctCalc', denom = 'Pitches')

#in play percentage
alphasinPlay_pct <- lapply(PitsampleSize, alphaCalc, metricFUN = 'inPlay_pctCalc', denom = 'Pitches')

#hit by pitch percentage
alphashbp_pct <- lapply(PitsampleSize, alphaCalc, metricFUN = 'hbp_pctCalc', denom = 'Pitches')

#similar to the plate appearance statistics, we will create a dataframe of the above six statistics to make plotting with ggplot easier

Pitch_alphas <- do.call(rbind, Map(data.frame, sampleSize = PitsampleSize, swStrAlpha = alphasswStr_pct, cStrAlpha = alphascStr_pct, foulAlpha = alphasfoul_pct, ballAlpha = alphasball_pct, inPlayAlpha = alphasinPlay_pct, hbpAlpha = alphashbp_pct))

```


Now that we have calculated our stabilization rates for each of our statistics, we can plot the data
```{r}

#first we will plot the plate appearance statistics
plot1 <- PA_alphas %>% ggplot() +
  geom_line(aes(sampleSize, pitPerPAAlpha, color = 'Pitches Per PA', group = 1, text = paste("Sample Size: ", sampleSize, "<br>Alpha Value: ", round(pitPerPAAlpha, 3), "<br>Statistic: ", 'Pitches Per PA'))) +
  geom_line(aes(sampleSize, K_pctAlpha, color = 'Strikeout Rate', group = 1, text = paste("Sample Size: ", sampleSize, "<br>Alpha Value: ", round(K_pctAlpha, 3), "<br>Statistic: ", 'Strikeout Rate'))) +
  theme_stata() +
  labs(title = 'Stabilization Rates for Pitcher Statistics by Number of PAs', subtitle = 'Statistics stabilize at an alpha of 0.7 (Horizontal Line)', x = 'Sample Size', y = 'Alpha Level') +
  geom_hline(yintercept = 0.7) + theme(legend.background = element_rect(fill="lightgreen", size=0.5, linetype="solid"), legend.title = element_blank())

ggplotly(plot1, tooltip = "text")


```

The graph shows that strikeout rate stabilizes faster than pitches per plate appearance

```{r}

#next we will plot the per-pitch statistics
plot2 <- Pitch_alphas %>% ggplot() +
  geom_line(aes(sampleSize, swStrAlpha, color = 'Swinging Strike Pct', group = 1, text = paste("Sample Size: ", sampleSize, "<br>Alpha Value: ", round(swStrAlpha, 3), "<br>Statistic: ", 'Swinging Strike Pct'))) +
  geom_line(aes(sampleSize, cStrAlpha, color = 'Called Strike Pct', group = 1, text = paste("Sample Size: ", sampleSize, "<br>Alpha Value: ", round(cStrAlpha, 3), "<br>Statistic: ", 'Called Strike Pct'))) +
  geom_line(aes(sampleSize, foulAlpha, color = 'Foul Pct', group = 1, text = paste("Sample Size: ", sampleSize, "<br>Alpha Value: ", round(foulAlpha, 3), "<br>Statistic: ", 'Foul Pct'))) +
  geom_line(aes(sampleSize, ballAlpha, color = 'Ball Pct', group = 1, text = paste("Sample Size: ", sampleSize, "<br>Alpha Value: ", round(ballAlpha, 3), "<br>Statistic: ", 'Ball Pct'))) +
  geom_line(aes(sampleSize, inPlayAlpha, color = 'In Play Pct', group = 1, text = paste("Sample Size: ", sampleSize, "<br>Alpha Value: ", round(inPlayAlpha, 3), "<br>Statistic: ", 'In Play Pct'))) +
  geom_line(aes(sampleSize, hbpAlpha, color = 'Hit By Pitch Pct', group = 1, text = paste("Sample Size: ", sampleSize, "<br>Alpha Value: ", round(hbpAlpha, 3), "<br>Statistic: ", 'Hit By Pitch Pct'))) +
  theme_stata() +
  labs(title = 'Stabilization Rates for Pitcher Statistics by Number of Pitches', subtitle = 'Statistics stabilize at an alpha of 0.7 (Horizontal Line)', x = 'Sample Size', y = 'Alpha Level') +
  geom_hline(yintercept = 0.7) + theme(legend.background = element_rect(fill="lightgreen", size=0.5, linetype="solid"), legend.title = element_blank())

ggplotly(plot2, tooltip = "text")


```
The graph shows that swinging strike rate stabilizes the fastest, followed by in play percentage
Hit by pitch rate never stabilizes, suggesting it is not a reliable stat that pitchers can control. It is more based on randomness than skill.
